A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-09 tex/codatadp.tex
Another refactor to solution soundness, now work with lazy construction of valuation.

We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
interprets wrongly-applied selectors based on equality information in $\Fc$,
interprets the variables and $x$ of $\Fc$ based on a mapping $\Val$ from the equivalence classes of $\Fc$ to values, 
which is constructed based on the following rules.

\rn{Ord}
\(
\inferrule{
  \ec{x} \not\in \vdom(\Val)
  \quad
  x : \tau
  \quad
  \tau \in \Nondata
  \quad
  c \not\in \vrange(\Val)
}{
  \Val := \Val \cup \ec{x} \mapsto c
}
\)

@@

\rn{D-Base}
\(
\inferrule{
  \ec{x} \not\in \vdom(\Val)
  \quad
  x : \tau
  \quad
  \tau \in \Data
  \quad
  \ec{x} \text{ ctr-free}
  \quad
  c \not\in \vrange(\Val)
}{
  \Val := \Val \cup \ec{x} \mapsto c
}
\)

We assume these rules are applied in the order in which they are written, for instance, 
the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.

@@

If $\tau \in \Data$,
we first show that $\M( t ) = \Val \ec{x}$ for all non-selector terms $t$ that occur in datatype equivalence classes $\ec{x}$.
We show this by induction on the size of $\Val \ec{x}$.
In the case where $t$ is a variable, then $\M( t ) = \Val \ec{ t } = \Val \ec{ x }$.
In the case where $t$ is a constructor term $C( \vec u )$, 
then $\Val \ec{x}$ must have been assigned by \rn{D-Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.
Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val \ec{ x }$.

>>>>

We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
interprets wrongly-applied selectors based on equality information in $\Fc$,
interprets the variables and $x$ of $\Fc$ based on a mapping $\Val$ from the equivalence classes of $\Fc$ to values.
Initially, $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Fc$,
where $v_{\ec{x}}$ is a fresh variable of the same type as $x$ not occurring in $\Fc$.
Then, $\Val$ is updated according to the following rules:

\begin{comment}
\rn{Ord}
\(
\inferrule{
  \ec{x} \not\in \vdom(\Val)
  \quad
  x : \tau
  \quad
  \tau \in \Nondata
  \quad
  c \not\in \vrange(\Val)
}{
  \Val := \Val \cup \ec{x} \mapsto c
}
\)

@@

\rn{D-Base}
\(
\inferrule{
  \ec{x} \not\in \vdom(\Val)
  \quad
  x : \tau
  \quad
  \tau \in \Data
  \quad
  \ec{x} \text{ ctr-free}
  \quad
  c \not\in \vrange(\Val)
}{
  \Val := \Val \cup \ec{x} \mapsto c
}
\)
\end{comment}

\rn{Ind}
\(
\inferrule{
  %x : \tau
  %\quad
  %\tau \in \Codata
  v_{\ec{x}} \in FV( \Val )
  \quad
  C( t_1, \ldots, t_n ) \in \ec{x}
  \quad
  C \in \Ctr
}{
  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
}
\)

\rn{Base}
\(
\inferrule{
  %x : \tau
  %\quad
  %\tau \in \Codata
  %\quad
  \ec{x} \mapsto v_{\ec{x}} \in \Val
  %\quad
  %\ec{x} \text{ ctr-free}
  \quad
  c \not\sim c' \text{ for all } c' \in \vrange(\Val)
}{
  \Val := \Val \{ v_{\ec{x}} \mapsto c \}
}
\)

\rem{define $\sim$}
We assume these rules are applied to completion in the order in which they are written,
that is, \rn{Ind} is applied exhaustively before applying \rn{Base}.
%the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
In the rule \rn{Ind}, we write $FV( \Val )$ to denote the free variables occurring in $\Val$.
Notice that this excludes variables beneath $\mu$-binders.
It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.

@@

If $\tau \in \Data$,
we first show that $\M( t ) = \Val \ec{x}$ for all non-selector terms $t$ that occur in datatype equivalence classes $\ec{x}$.
We show this by induction on the size of $\Val \ec{x}$.
In the case where $t$ is a variable, then $\M( t ) = \Val \ec{ t } = \Val \ec{ x }$.
In the case where $t$ is a constructor term $C( \vec u )$, 
then $\Val \ec{x}$ must have been assigned by \rn{Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.
Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val \ec{ x }$.
